Relationships in the extremes and their influence on competition and coexistence: Supporting information

### 2021-06-29 paper/SupMat.Rnw
fixed the use of \hat in several places to make the latexc compile

\begin{document}
\SweaveOpts{concordance=TRUE}
%\SweaveOpts{concordance=TRUE}

@@

We begin by decomposing 
the invasion growth rate,
\begin{equation}
\olr_{i \bs i}=\E[r_i(E_i,C_{\i \bs i})].
\end{equation}
The invasion growth rate with both $E_i$ and $C_{i \bs i}$ constant at their means is 
\begin{align}
\epsilon_i^0 = r_i(\overline{E}_i,\overline{C}_{i \bs i}).
\end{align}
The additional contribution of variation in $E_i$ to the invasion growth rate is then
\begin{equation}
\olep_i^E = \E[r_i(E_i,\overline{C}_{i \bs i})]-\epsilon_i^0,
\end{equation}
whereas the additional contribution of variation in $C_{i \bs i}$ is 
\begin{equation}
\olep_i^C=\E[r_i(\overline{E}_i,C_{i \bs i})]-\epsilon_i^0.
\end{equation}
We then define
\begin{align}
\olep_i^{EC} &= \E[r_i(E_i,C_{i \bs i})]-\olep_i^E-\olep_i^C-\epsilon_i^0 \\
&= \olr_{i \bs i} - \olep_i^E - \olep_i^C - \epsilon_i^0.
\end{align}
Thus we have the primary decomposition 
\begin{equation}
\olr_{i \bs i} = \epsilon_i^0 + \olep_i^E + \olep_i^C + \olep_i^{EC}.
\end{equation}
But some of $\olep_i^{EC}$ comes from the fact that $E_i$ and $C_{i \bs i}$ are
both varying, and some comes from covariation in these two quantities. So we next separate these
contributions in a secondary decomposition of $\olep_i^{EC}$.

@@

\item Estimate $\bar r_1$ for the left-tail associated case as $\hat \bar r_1 = \mean_i (\ln [1-\delta+\delta\exp(b_{l1}^{(i)}-b_{l2}^{(i)})])$ where $b_{l1}^{(i)} = \sigma \tilde b_{l1}^{(i)} + \mu_1$ and $b_{l2}^{(i)} = \sigma \tilde b_{l2}^{(i)} + \mu_2$. The standard error of this estimate is $\se (\hat \bar r_1) =\frac{\sd _i (\ln[1-\delta +\delta \exp(b_{l1}^{(i)}-b_{l2}^{(i)})])}{\sqrt M}$, where $\sd _i(\cdot)$ is standard deviation.

\item Estiate $\bar r_1$ for the left-tail associated case as $\hat \bar r_1 = \mean_i (\ln [1-\delta+\delta\exp(b_{r1}^{(i)}-b_{r2}^{(i)})])$ where $b_{r1}^{(i)} = \sigma \tilde b_{r1}^{(i)} + \mu_1$ and $b_{r2}^{(i)} = \sigma \tilde b_{r2}^{(i)} + \mu_2$. The standard error of this estimate is $\se (\hat \bar r_1) =\frac{\sd _i (\ln[1-\delta +\delta \exp(b_{r1}^{(i)}-b_{r2}^{(i)})])}{\sqrt M}$.

\item Estimate $\bar r_1$ in the symmetric tail association case as $\hat \bar r_1=\mean_i(\ln[1-\delta+\delta \exp(u_i)])$, where $u_i = \sqrt{(2\sigma^2-2\rho)}\tilde u_i + \mu_1 - \mu_2$. The standard error of this estimation is $\se(\hat \bar r_1) = \frac {\sd_i(\ln[1-\delta+\delta \exp(u_i)])}{\sqrt M}$.

\item Estimate $\bar r_1^{\#}$ (for all three tail-assocaition cases) as $\hat \bar r_1^{\#} = \mean_i(\ln[1-\delta+\delta \exp(u_i)])$, where $u_i = \sqrt{2\sigma^2}\tilde u_i + \mu_1 - \mu_2$. The standard error is $\se(\hat \bar r_1^{\#}) = \sd_i \frac{ln[1-\delta+\delta \exp(u_i)]}{\sqrt M}$.

\item Estimate $\bar r_2^{\#}$ (for all three tail-assocaition cases) as $\hat \bar r_2^{\#} = \mean_i(\ln[1-\delta+\delta \exp(u_i)])$, where $u_i = \sqrt{2\sigma^2}\tilde u_i$. The standard error is $\se(\hat \bar r_2^{\#}) = \sd_i \frac{ln[1-\delta+\delta \exp(u_i)]}{\sqrt M}$.

\item $\hat {\Delta I} = \hat \bar r_1 - \hat \bar r_1^{\#} + \hat \bar r_2^{\#}$, estimated separately for the left-tail associated, right-tail associated and symmetric noise cases. We obtained conservative standard error estimates by adding the standard error of the estimates of $\hat \bar r_1, \hat \bar r_1^{\#}$, and $\hat \bar r_2^{\#}$.

>>>>

\begin{document}
%\SweaveOpts{concordance=TRUE}

@@

We begin by decomposing 
the invasion growth rate,
\begin{equation}
\olr_{i \bs i}=\E[r_i(E_i,C_{i \bs i})].
\end{equation}
The invasion growth rate with both $E_i$ and $C_{i \bs i}$ constant at their means is 
\begin{align}
\epsilon_i^0 = r_i(\overline{E}_i,\overline{C}_{i \bs i}).
\end{align}
The additional contribution of variation in $E_i$ to the invasion growth rate is then
\begin{equation}
\olep_i^E = \E[r_i(E_i,\overline{C}_{i \bs i})]-\epsilon_i^0,
\end{equation}
whereas the additional contribution of variation in $C_{i \bs i}$ is 
\begin{equation}
\olep_i^C=\E[r_i(\overline{E}_i,C_{i \bs i})]-\epsilon_i^0.
\end{equation}
We then define
\begin{align}
\olep_i^{EC} &= \E[r_i(E_i,C_{i \bs i})]-\olep_i^E-\olep_i^C-\epsilon_i^0 \\
&= \olr_{i \bs i} - \olep_i^E - \olep_i^C - \epsilon_i^0.
\end{align}
Thus we have the primary decomposition 
\begin{equation}
\olr_{i \bs i} = \epsilon_i^0 + \olep_i^E + \olep_i^C + \olep_i^{EC}.
\end{equation}
But some of $\olep_i^{EC}$ comes from the fact that $E_i$ and $C_{i \bs i}$ are
both varying, and some comes from covariation in these two quantities. So we next separate these
contributions in a secondary decomposition of $\olep_i^{EC}$.

@@

\item Estimate $\bar r_1$ for the left-tail associated case as $\hat{\bar r_1} = \mean_i (\ln [1-\delta+\delta\exp(b_{l1}^{(i)}-b_{l2}^{(i)})])$ where $b_{l1}^{(i)} = \sigma \tilde b_{l1}^{(i)} + \mu_1$ and $b_{l2}^{(i)} = \sigma \tilde b_{l2}^{(i)} + \mu_2$. The standard error of this estimate is $\se (\hat{\bar r_1}) =\frac{\sd _i (\ln[1-\delta +\delta \exp(b_{l1}^{(i)}-b_{l2}^{(i)})])}{\sqrt M}$, where $\sd _i(\cdot)$ is standard deviation.

\item Estiate $\bar r_1$ for the left-tail associated case as $\hat{\bar r_1} = \mean_i (\ln [1-\delta+\delta\exp(b_{r1}^{(i)}-b_{r2}^{(i)})])$ where $b_{r1}^{(i)} = \sigma \tilde b_{r1}^{(i)} + \mu_1$ and $b_{r2}^{(i)} = \sigma \tilde b_{r2}^{(i)} + \mu_2$. The standard error of this estimate is $\se (\hat{\bar r_1}) =\frac{\sd _i (\ln[1-\delta +\delta \exp(b_{r1}^{(i)}-b_{r2}^{(i)})])}{\sqrt M}$.

\item Estimate $\bar r_1$ in the symmetric tail association case as $\hat{\bar r_1}=\mean_i(\ln[1-\delta+\delta \exp(u_i)])$, where $u_i = \sqrt{(2\sigma^2-2\rho)}\tilde u_i + \mu_1 - \mu_2$. The standard error of this estimation is $\se(\hat{\bar r_1}) = \frac {\sd_i(\ln[1-\delta+\delta \exp(u_i)])}{\sqrt M}$.

\item Estimate $\bar r_1^{\#}$ (for all three tail-assocaition cases) as $\hat{\bar r_1^{\#}} = \mean_i(\ln[1-\delta+\delta \exp(u_i)])$, where $u_i = \sqrt{2\sigma^2}\tilde u_i + \mu_1 - \mu_2$. The standard error is $\se(\hat{\bar r_1^{\#}}) = \sd_i \frac{ln[1-\delta+\delta \exp(u_i)]}{\sqrt M}$.

\item Estimate $\bar r_2^{\#}$ (for all three tail-assocaition cases) as $\hat{\bar r_2^{\#}} = \mean_i(\ln[1-\delta+\delta \exp(u_i)])$, where $u_i = \sqrt{2\sigma^2}\tilde u_i$. The standard error is $\se(\hat{\bar r_2^{\#}}) = \sd_i \frac{ln[1-\delta+\delta \exp(u_i)]}{\sqrt M}$.

\item $\hat {\Delta I} = \hat{\bar r_1} - \hat{\bar r_1^{\#}} + \hat{\bar r_2^{\#}}$, estimated separately for the left-tail associated, right-tail associated and symmetric noise cases. We obtained conservative standard error estimates by adding the standard error of the estimates of $\hat{\bar r_1}, \hat{\bar r_1^{\#}}$, and $\hat{\bar r_2^{\#}}$.
